Executive Constraints and Civil Conflict Onset

### 2020-01-29 paper.Rmd
Begin work on results section

---
title: Executive Constraints and Civil Conflict Onset
author: J.L.A. Krusell
date: "`r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
csl: assets/american-political-science-association.csl
link-citations: true
documentclass: scrreprt
fontsize: 11pt
linestretch: 1.5
header-includes:
    - \usepackage{amsmath}
    - \usepackage{etoolbox}
    - \usepackage{lualatex-math}
    - \usepackage{unicode-math}
    - \appto\Shaded{\singlespacing}
    - \BeforeBeginEnvironment{align*}{\begin{singlespace}\vspace*{-\baselineskip}}
    - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
output:
    pdf_document:
        latex_engine: lualatex
        toc: true
        toc_depth: 2
        pandoc_args: ["--syntax-definition", "assets/stan.xml",
                      "--highlight-style", "kate"]
    html_document:
        smart: true
        toc: true
abstract: Lorem ipsum
---

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
knitr::opts_chunk$set(echo = F, eval = T)

library(dplyr)
library(extraDistr)
library(ggplot2)
library(gridExtra)
library(thesis.utils)
```

@@

To examine the relationship between civil conflict onset and executive
constraints, I construct a dataset of relevant variables and run a
bayesian structural equation model with a measurement model estimating
executive constraints as a latent construct and a multilevel logistic
regression relating the latent construct to the observed outcome,
civil conflict. The use of bayesian methods is particularly
advantageous in this setting since the constructed dataset is
comprised of the entire finite population of countries within a
non-random subset of years given only non-stochastic
missingness. Traditional frequentist estimators are inappropriate due
to the assumption that the observed data is drawn from a random,
repeatable generative mechanism. This is readily apparent when
attempting to interpret a frequentist confidence interval in the given
context. A standard confidence interval for a selected level of
significance, $\alpha$, reports only that $(1 - \alpha) 100 \%$ of the
calculated confidence intervals from additional inferences done on
repeated samples are expected to contain the true population value
[@Western1994]. However, for the current analysis the desired target
is really to be able to calculate directly the probability that the
current uncertainty interval contains the true value given the data
and the model specification. Bayesianism offers a solution by casting
probability as degrees of belief in some event wherein inference is
then the process of updating those prior beliefs in light of new
data. When using bayesian methods, parameters such as regression
coefficients are considered to be random, meaning that prior beliefs
in the form of probability distributions can be assigned and posterior
probability distributions estimated from observed data [@Western1999].

@@

[^model]: The model itself is written in Stan [@Carpenter2017] and run
    using the `CmdStan` interface [@Team2019a] with
    pre/post-processing done using `R` [@Team2019]. The model code is
    available in [Appendix A](#appendix-a) and a fully reproducible
    pipeline composed of all replication scripts is available at
    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).

@@

I model civil conflict onset as a bayesian multilevel regression where
onset is a function of a set of predictors, including executive
constraints and an array of control variables, together with several
grouping factors included as partially pooled varying intercepts. The
units of analysis are all independent country-years from 1946 until
2016, as determined by data availability, with *I* countries indexed
by $i = 1 \dotsc I$ and *T* years indexed by $t = 1 \dotsc T$. The
dependent variable, conflict onset, is a Bernoulli distributed binary
variable where 1 denotes the onset of at least one new civil conflict
in a given country-year. The chosen link function transforming the
linear regression output into the probability of conflict onset,
$\nu_{it}$, is the inverse logistic such that $\nu_{it} \in [0,
1]$. [^invlogit]

@@

where **X** is a single row matrix of predictors, $\beta$ a vector of
regression slopes, and $\zeta^k$ the set of varying intercepts for *K*
grouping factors indexed by $j_k$. For the present analysis there is
only two grouping factors, year and country, to account for
cross-sectional unmodeled heterogeneity between countries and over
time [@Shor2007]. All predictors are also lagged by one year to
control for time endogeneity and standardized by substracting the mean
and dividing by the standard deviation for each variable to aid in
computation as well as comparison of the final $\beta$ estimates.

@@

Finally, for the varying intercepts modeling group-specific
variability, I use hierarchical priors to provide a regularizing
effect against overfitting [@Gelman2007].
\begin{align*}
    \zeta^k_j & \sim \mathsf{Normal}(0, \sigma^k) \\
    \sigma^k & \sim \mathsf{HalfCauchy}(0, 1)
\end{align*}

@@

The dependent variable, *civil conflict onset*, is taken from the
UCDP/PRIO Armed Conflict Dataset v19.1 [@Gleditsch2002;
@Pettersson2019] where intra-state conflict is defined as a violent
contest over a government or territory between a state and non-state
actor with a minimum threshold of 25 battle deaths within a single
country-year with a coverage from 1946 to 2018.  The data was merged
based on country location for conflicts labeled as internal or
internationalized internal --- *i.e.* where *type_of_conflict* was
either 3 or 4. Multiple conflicts within the same years were
aggregated to form a binary variable coded as 1 to denote the onset of
at least one new civil conflict and 0 for all other years, including
years of ongoing conflict.[^conflict_type]

@@

To measure executive constraints I use data from the Varieties of
Democracy (V-Dem) dataset v9.0 [@Coppedge2019]. V-Dem is a
cross-national time series dataset providing multidimensional,
disaggregated data on democracy and political institutions for over
200 countries with an overall year coverage from 1789 until 2018. The
principal data is drawn from expert coded surveys involving a global
network of over 3000 coders, most of whom are academics as well as
professionals working in policy related fields. Individual ratings are
aggregated into latent variable estimates using hierarchical ordinal
item response theory (IRT) models taking into account rater
reliability and varying rater thresholds
[@Pemstein2019]. [^thresholds] The result is over 220 lower level
indicators that are also combined through various aggregation
techniques to produce over 80 indices. Compared to datasets such as
Polity IV, which has traditionally been used in most analyses of civil
conflict, V-Dem provides much more fine grained, nuanced data, and
most importantly, includes measures of uncertainty propagated from the
IRT models. This latter point is crucial since ignoring measurement
error, especially in models with noisy data and small effect sizes as
is often the case in conflict studies, can bias estimates and lead to
overconfident results [@Treier2008; @Marquardt2019].

@@

For the purposes of this paper, I estimate an index measuring
executive constraints from a factor analysis of several manifest
variables drawn from the V-Dem dataset. The aim is to capture the
extent to which the executive is *de facto* constrained by other
government institutions --- more specifically, the legislature,
judiciary, and independent investigative bodies --- rather than the
*de jure* powers of any one institution *vis-à-vis* the executive, and
to do so using a latent variable model that avoids the arbitrary
weighting of linear additive indices.[^additive] The aggregation
method is largely similar to the accountability index created by
@Luhrmann2017 and the presidentialism index created by @Sigman2018. By
creating a separate index, however, I am able to adjust the list of
included manifest variables as well as embed the factor analysis
within the larger regression model such that the uncertainty from the
latent factor propagates as a result of the estimation process.

@@

I first model the measurement error in the manifest variables by
assuming that the observed values are drawn from a normal distribution
where the location parameters are the latent true values and the
scale, $u$, the observed posterior standard deviations derived from
the IRT posterior distributions. [^err]
\[ Z_{ij} \sim \mathsf{Normal}(Z^\ast_{ij}, u_{ij}) \]

@@

The latent values for the manifest variables are then modeled as
functions of the latent factor $\theta$ where the mapping function is
a standard factor analysis.
\[ Z^\ast_{ij} \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]

where $\lambda$ is the manifest variable specific factor loading,
$\gamma$ the population intercept, and $\psi$ the scale of the error
term. I assume local item independence such that the error terms,
$(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
Novick). Furthermore, for the purposes of identifying the likelihood,
the scale of the latent factor, $\theta$, is fixed to mean $0$ and
standard deviation $1$ while the factor loadings, $(\lambda_1, \dotsc,
\lambda_n)$, are constrained to be positive [@Jackman2009]. The
positive constraints follow theory since all of the manifest variables
exhibit the same scale direction and should be positively correlated
with the latent factor.[^corr] Finally, similar to the regression
coefficients, I select weakly informative priors.
\begin{align*}
    \gamma & \sim \mathsf{Normal}(0, 5) \\
    \lambda & \sim \mathsf{LogNormal}(0, 1) \\
    \psi & \sim \mathsf{Weibull}(2, 1)
\end{align*}

@@

### Additional Variables

@@

In addition to state capacity, I also include a number of standard
variables drawn largely from previous literature. The list of
variables are not meant to provide an all encompassing theory of civil
conflict, but rather roughly capture the most relevant dimensions of
opportunity structure for armed conflict.  This includes a single year
lagged measure of GDP per capita growth from PWT in order to capture
short term economic shocks, which may affect both the strength and
legitimacy of the state [@Bazzi2014]. I also include a measure of mean
elevation from the GROWup dataset [@Girardin2015] to capture terrain
advantageous for conducting an insurgency [@Fearon2003] as well as a
measure of population density derived from population data from PWT
and the area in square kilometers for each country from
GROWup. Countries with a higher population density may require a more
extensive and competent central administration and policing force in
order to suppress violent dissent [@Fearon2003]. There's also the
possibility that a high population density increases the competition
for resources. If the state is identified as the source of a
particular scarcity, the risk of civil conflict may increase. Finally,
I include measures of ethnic fractionalization in the form of the
number of political relevant ethnic groups and the percent of the
population facing discrimination. Where there are significant ethnic
cleavages and state sponsored discrimination, ethnic ties can
facilitate collective action thereby lowering the mobilization costs
of rebel groups [@Cederman2011].

To control for the time endogeneity of conflict --- past and current
civil conflicts can impact the present risk of conflict onset as well
as the level of executive constraints --- I include a dummy variable
for ongoing conflict and the log of peace years [@Hegre2015a]. [^time]
To account for spillover effects, I include a dummy variable for
whether a neighbouring country is currently experiencing a civil
conflict. One notable lack is a general measure of political
instability, which I omit since it would be highly correlated with
executive constraints and could potentially bias results, at least for
years of crisis where the risk of civil conflict is highest. I do,
however, include a dummy variable for independence year since newly
sovereign states may be particularly weak.

@@

    Regarding the peace years, the variable is calculated as
    the number of years since the last conflict-year, including
    conflict-years from extrasystemic conflicts. For countries left
    censored due to the dataset starting date, I set the start of the
    peace year calculation to 1946, the end of World War II. The
    rationale is that World War II was a significant international
    event that reshaped the world order and even had domestic
    consequences for countries not directly involved. For newly
    independent countries, I use the independence date as the start
    year.

@@

### Latent Factor Model

### Conflict Onset Regression

```{r eval = F}
beta <- readRDS("posteriors/summary/beta.rds")
plot_pars(beta)
```

```{r eval = F}
input.df <- readRDS("posteriors/input_data.rds")
p_hat <- readRDS("posteriors/summary/predicted_probs.rds")

@@

```{r eval = F}
p <- apply(p_hat, 2, median)
x <- evalmod(scores = p, labels = input.df$lepisode_onset)
plot(x, curvetype = c("ROC", "PRC"))
```

```{r eval = F}
# Interaction plot
intercept <- readRDS("posteriors/summary/intercept.rds")
theta <- readRDS("posteriors/summary/theta.rds") %>% apply(2, median) %>% sort

state <- quantile(input.df$cgdppc, probs = c(0.2, .5, 0.8))

>>>>

---
title: Executive Constraints and Civil Conflict Onset
author: J.L.A. Krusell
date: "`r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
csl: assets/american-political-science-association.csl
link-citations: true
documentclass: scrreprt
fontsize: 11pt
linestretch: 1.5
header-includes:
    - \usepackage{amsmath}
    - \usepackage{etoolbox}
    - \usepackage{lualatex-math}
    - \usepackage{unicode-math}
    - \appto\Shaded{\singlespacing}
    - \BeforeBeginEnvironment{align*}{\begin{singlespace}\vspace*{-\baselineskip}}
    - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
    - \counterwithout{table}{chapter}
    - \counterwithout{figure}{chapter}
output:
    pdf_document:
        fig_caption: true
        keep_tex: true
        latex_engine: lualatex
        toc: true
        toc_depth: 2
        pandoc_args: ["--syntax-definition", "assets/stan.xml",
                      "--highlight-style", "kate",
                      "--filter", "pandoc-crossref"]
    html_document:
        smart: true
        toc: true
abstract: Lorem ipsum
---

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
library(dplyr)
library(extraDistr)
library(ggplot2)
library(gridExtra)
library(magrittr)
library(kableExtra)
library(knitr)
library(tidyr)

# Load last to ensure nothing overwrites our namespace
library(thesis.utils)

options(knitr.kable.NA = ".")
opts_chunk$set(echo = F, eval = T, warning = F, fig.width = 5,
               fig.height = 3)

@@

To examine the relationship between civil conflict onset and executive
constraints, I construct a dataset of relevant variables and run a
bayesian structural equation model with a measurement portion estimating
executive constraints as a latent construct and a multilevel logistic
regression relating the latent construct to the observed outcome,
civil conflict. The use of bayesian methods is particularly
advantageous in this setting since the constructed dataset is
comprised of the entire finite population of countries within a
non-random subset of years given only non-stochastic
missingness. Traditional frequentist estimators are inappropriate due
to the assumption that the observed data is drawn from a random,
repeatable generative mechanism. This is readily apparent when
attempting to interpret a frequentist confidence interval in the present
context. A standard confidence interval for a selected level of
significance, $\alpha$, reports only that $(1 - \alpha) 100 \%$ of the
calculated confidence intervals from additional inferences done on
repeated samples are expected to contain the true population value
[@Western1994]. However, for the current analysis the desired target
is really to be able to calculate directly the probability that the
current uncertainty interval contains the true value given the data
and the model specification. Bayesianism offers a solution by casting
probability as degrees of belief in some event wherein inference is
then the process of updating those prior beliefs in light of new
data. When using bayesian methods, parameters such as regression
coefficients are considered to be random, meaning that prior beliefs
in the form of probability distributions can be assigned and posterior
probability distributions estimated from observed data [@Western1999].

@@

[^model]: The model code is written in Stan [@Carpenter2017] and was
    run with the `CmdStan` interface [@Team2019a] using four parallel
    chains at 2000 iterations each with a warmup of 1000. Convergence
    was assessed according to the potential scale reduction factor
    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well as
    standard Stan diagnostics. The full code is available in
    [Appendix A](#appendix-a) and a fully reproducible pipeline
    compromised of replication scripts is available at
    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).

@@

I model civil conflict onset as a multilevel regression where onset is
a function of a set of predictors, including executive constraints and
an array of control variables, together with several grouping factors
included as partially pooled varying intercepts. The units of analysis
are all independent country-years from 1951 until 2017, as determined
by data availability, with *I* countries indexed by $i = 1 \dotsc I$
and *T* years indexed by $t = 1 \dotsc T$. The dependent variable,
conflict onset, is a Bernoulli distributed binary variable where 1
denotes the onset of at least one new civil conflict in a given
country-year. The chosen link function transforming the linear
regression output into the probability of conflict onset, $\nu_{it}$,
is the inverse logistic such that $\nu_{it} \in [0, 1]$. [^invlogit]

@@

where **X** is a vector of predictors, $\beta$ the vector of
regression slopes, and $\zeta^k$ the set of varying intercepts for *K*
grouping factors indexed by $j_k$. For the present analysis there are
only two grouping factors, year and country, to account for
cross-sectional unmodeled heterogeneity between countries and over
time [@Shor2007]. All predictors are also lagged by one year to
control for time endogeneity and standardized by substracting the mean
and dividing by the standard deviation for each variable to aid in
computation as well as comparison of the final $\beta$ estimates.

@@

For the varying intercepts modeling group-specific variability, I use
hierarchical priors to provide a regularizing effect against
overfitting [@Gelman2007].
\begin{align*}
    \zeta^k_j & \sim \mathsf{Normal}(0, \sigma^k) \\
    \sigma^k & \sim \mathsf{HalfCauchy}(0, 1)
\end{align*}

@@

The dependent variable, *civil conflict onset*, is taken from the
UCDP/PRIO Armed Conflict Dataset v19.1 [@Gleditsch2002;
@Pettersson2019] where intra-state conflict is defined as a violent
contest over a government or territory between a state and non-state
actor with a minimum threshold of 25 battle deaths within a single
country-year with a coverage from 1946 to 2018.  The data was merged
based on the government side for conflicts labeled as internal or
internationalized internal --- *i.e.* where *type_of_conflict* was
either 3 or 4. Multiple conflicts within the same years were
aggregated to form a binary variable coded as 1 to denote the onset of
at least one new civil conflict and 0 for all other years, including
years of ongoing conflict.[^conflict_type]

@@

To measure executive constraints I use data from the Varieties of
Democracy (V-Dem) dataset v9.0 [@Coppedge2019]. V-Dem is a
cross-national time series dataset providing multidimensional,
disaggregated data on democracy and political institutions for over
200 countries with an overall year coverage from 1789 until 2018. The
principal data is drawn from expert coded surveys involving a global
network of over 3000 coders, most of whom are academics or
professionals working in policy related fields. Individual ratings are
aggregated into latent variable estimates using hierarchical ordinal
item response theory (IRT) models which take into account rater
reliability and varying rater thresholds
[@Pemstein2019]. [^thresholds] The result is over 220 lower level
indicators that are also combined through various aggregation
techniques to produce over 80 indices. Compared to datasets such as
Polity IV, which has traditionally been used in most analyses of civil
conflict, V-Dem provides much more fine grained, nuanced data, and
most importantly, includes measures of uncertainty propagated from the
IRT models. This latter point is crucial since ignoring measurement
error, especially in models with noisy data and small effect sizes as
is often the case in conflict studies, can bias estimates and lead to
overconfident results [@Treier2008; @Marquardt2019].

@@

For the purposes of this paper I estimate an index measuring
executive constraints from a factor analysis of multiple manifest
variables drawn from the V-Dem dataset. The aim is to capture the
extent to which the executive is *de facto* constrained by other
government institutions --- more specifically, the legislature,
judiciary, and independent investigative bodies --- rather than the
*de jure* powers of any one institution *vis-à-vis* the executive, and
to do so using a latent variable aggregation that avoids the arbitrary
weighting of linear additive indices.[^additive] The aggregation
method is largely similar to the accountability index created by
@Luhrmann2017 and the presidentialism index created by @Sigman2018. By
creating a separate index, however, I am able to adjust the list of
included manifest variables as well as embed the factor analysis
within the larger regression model such that the uncertainty from the
latent factor propagates as a result of the estimation process.

@@

I first model the measurement error in the manifest variables by
assuming that the observed values are drawn from a normal distribution
where the location parameters are the unobserved true values and the
scale, $u$, the observed posterior standard deviations derived from
the IRT posterior distributions. [^err]
\[ Z_{ij} \sim \mathsf{Normal}(Z^\ast_{ij}, u_{ij}) \]

@@

The latent values for the manifest variables are then modeled as
functions of the latent factor $\theta$ with a standard factor
analytic mapping function.
\[ Z^\ast_{ij} \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]

In the preceding equation, $\lambda$ is the variable specific factor
loading, $\gamma$ the population intercepts, and $\psi$ the scale of
the error term. I assume local item independence such that the error
terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
Novick). Furthermore, for the purposes of identifying the likelihood,
the scale of the latent factor, $\theta$, is fixed to mean $0$ and
standard deviation $1$ while the factor loadings, $(\lambda_1, \dotsc,
\lambda_n)$, are constrained to be positive [@Jackman2009]. The
positive constraints follow theory since all of the manifest variables
exhibit the same scale direction and should be positively correlated
with the latent factor.[^corr] Finally, similar to the regression
coefficients, I select the following weakly informative priors:
\begin{align*}
    \gamma & \sim \mathsf{Normal}(0, 5) \\
    \lambda & \sim \mathsf{LogNormal}(0, 1) \\
    \psi & \sim \mathsf{Weibull}(2, 1)
\end{align*}

@@

### Additional Covariates

@@

In addition to state capacity, I also include a number of standard
variables drawn largely from previous literature. The list of
variables are not meant to provide an all encompassing theory of civil
conflict, but rather roughly capture the opportunity structures for
armed conflict.  This includes a single year lagged measure of GDP per
capita growth from PWT in order to capture short term economic shocks,
which may affect both the strength and legitimacy of the state
[@Bazzi2014]. I also include a measure of mean elevation from the
GROWup dataset [@Girardin2015] to capture terrain advantageous for
conducting an insurgency [@Fearon2003] as well as a measure of
population density derived from population data from PWT and the area
in square kilometers for each country from GROWup. Countries with a
higher population density may require a more extensive and competent
central administration and policing force in order to suppress violent
dissent [@Fearon2003]. There's also the possibility that a high
population density increases the competition for resources. If the
state is identified as the source of a particular scarcity, the risk
of civil conflict may increase. Finally, I include measures of ethnic
fractionalization in the form of the number of political relevant
ethnic groups and the percent of the population facing
discrimination. Where there are significant ethnic cleavages and state
sponsored discrimination, ethnic ties can facilitate collective action
thereby lowering the mobilization costs of rebel groups
[@Cederman2011].

To control for the time endogeneity of conflict --- past and current
civil conflicts can impact the present risk of conflict onset as well
as the current level of executive constraints --- I include a dummy
variable for ongoing conflict [^time] and the natural log of peace
years [^peace_yrs] [@Hegre2015a]. To account for spillover effects, I
include a dummy variable for whether a neighbouring country is
currently experiencing a civil conflict. One notable lack is a general
measure of political instability, which I omit since it could be
highly correlated with executive constraints and would potentially
bias results, at least for years of crisis where the risk of civil
conflict is highest. I do, however, include a dummy variable for the
independence year since newly sovereign states may be particularly
weak.

@@

[^peace_yrs]: Regarding the peace years, the variable is calculated as
    the number of years since the last conflict-year, including
    conflict-years from extrasystemic conflicts. For countries left
    censored due to the dataset starting date, I set the start of the
    peace year calculation to 1946, the end of World War II. The
    rationale is that World War II was a significant international
    event that reshaped the world order and even had domestic
    consequences for countries not directly involved. For newly
    independent countries, I use the independence date as the start
    year.

@@

The analysis of the results is split into two sections. First, the
parameter estimates from the executive constraints measurement model
will be presented followed by the parameters pertaining to the
logistic regression as well as a discussion on the overall model fit
to the outcome variable, conflict onset.

### Measurement Model

```{r  fig.cap = "\\label{fig:fpars} hello", warning = F}
load("./posteriors/full_model/input.RData")
fa <- read_post("./posteriors/full_model/fa_posteriors.csv.gz")

lg <- colnames(lg_mm$obs)
nonlg <- colnames(nonlg_mm$obs)

lambda <- extract(fa, pars = "lambda")
colnames(lambda) <- c(lg, nonlg)

psi <- extract(fa, pars = "psi")
colnames(psi) <- c(lg, nonlg)

gamma <- extract(fa, pars = "gamma")
colnames(gamma) <- c(lg, nonlg)

delta_eta <- extract(fa, pars = c("delta", "eta"))

p1 <- plot_pars(lambda)
p2 <- plot_pars(psi)
p3 <- plot_pars(gamma)
p4 <- plot_pars(delta_eta)

grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

Table \ref{tab:fa} and figure \ref{fig:fpars} summarise the posterior
draws for the measurement model parameters according to $80$ percent
quantile intervals. According to the results, the factor loadings,
$\lambda$, for each manifest variable are fairly comparable, with the
exception of the lower value for *v2lgfunds* ("legislature controls
its own resources") and to a lesser extent *v2lgqstexp* ("legislature
questions executive branch officials in practice"). Relative to the
other manifest variables, both items are less informative in inferring
the values of the latent factor --- a $\lambda$ of zero would indicate
that the latent factor does not explain any of the variation in the
manifest variable. As for the variable specific intercepts, $\gamma$,
they are largely clustered around zero as to be expected given that
the manifest variables were standardized prior to the model run. Where
deviations do occur, most notably with several of the legislative
variables, it is because of the measurement error component of the
model shifting the mean for a specific variable away from zero. For
example, after adjusting for measurement error, the mean for
$v2lgfunds$ is on average slightly higher than the remaining manifest
variables, hence the negative intercept.

```{r fa, fig.height = 6, fig.width = 6, warning = F}
fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "eta")) %>%
    post_summarise(probs = c(0.10, 0.9)) %>%
    mutate(parameter = sub("\\[\\d+\\]", "", parameter) %>% sprintf("$\\%s$", .),
           interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)),
           manifest = c(rep(c(lg, nonlg), 3),
                        rep("v2lgbicam", 2))) %>%
    select(-`10%`, -`90%`) %>%
    pivot_wider(id_cols = manifest, names_from = parameter, values_from = interval)

rownames(fa_pars) <- sprintf("\\textit{%s}", fa_pars$manifest)
fa_pars$manifest <- NULL

kable(fa_pars, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "Posterior credible intervals for measurement model parameters") %>%
    kable_styling(latex_options = c("striped", "scale_down"), full_width = F,
                  protect_latex = F)
```

The $\delta$ and $\eta$ parameters correspond to the slope and
intercept in the linear predictor for the presence of a legislature,
*v2lgbicam*. The effect of their inclusion can be seen in figure
\ref{fig:bicam}. The latent factor, $\theta$, follows the
directionality of the manifest variables, greater values indicate a
more constrained executive and lower values vice versa. The density
plots indicate that on average, for country-years where there is no
legislature, *i.e.* $v2lgbicam$ is zero, the latent factor is lower,
although still conditional on the values for the remaining manifest
variables. Where the legislature is present, $\theta$ is distributed
according to the $Normal(0, 1)$ prior.

```{r fig.width = 3, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for $\\theta$"}
theta <- extract(fa, "theta")
plot_dens(theta, groups = lgbicam)

The final portion of the measurement model is the measurement error
adjustment. As described in the research design section, in order to
incorporate the posterior uncertainty from the V-Dem IRT aggregation,
the observed values for the manifest variables, $Z$, were assumed to
be drawn from normal distributions where the means were the unobserved
true values, $Z^{\ast}$, and the standard deviations the observed
posterior standard deviations as reported by V-Dem. The practical
effect is that where the reported error is larger, the uncertainty in
$Z^{\ast}$ estimates increases and the mean is drawn towards the
regression line in the factor analysis. In other words, the model is
more skeptical of outliers in the manifest variables. This posterior
shrinkage can be seen in figure \ref{fig:fa-line}. Figure
\ref{fig:err} meanwhile plots the range of the upper and low bounds of
the $80$ percent posterior quantile interval for an example manifest
variable in $Z^{\ast}$ against the V-Dem standard deviations as well
as the difference between the estimated posterior mean and the
observed mean.

```{r fig.cap = "\\label{fig:fa-line} Error"}
err <- read_post("posteriors/full_model/err_posteriors.csv.gz")

nonlg_est <- extract(err, "nonlg_est") %>%
    post_summarise %>%
    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
    mutate(interval = codehigh - codelow,
           se = nonlg_mm$se[, 1],
           obs = nonlg_mm$obs[, 1],
           diff = obs - median)

idx <- sample(1:ncol(theta), 50)
theta_medians <- theta[, idx] %>% apply(2, median)

s <- seq(min(theta_medians) - 0.1, max(theta_medians) + 0.1, 0.01)

x <- data.frame(point = c(nonlg_est$obs[idx], nonlg_est$median[idx]),
                group = rep(1:length(idx), 2),
                factor = c(rep(1, length(idx)), rep(2, length(idx))) %>% as.factor,
                theta = rep(theta_medians, 2))

y <- data.frame(yhat = median(gamma[, 6]) + median(lambda[, 6]) * s,
                theta = s)

ggplot(x, aes(theta, point, group = group, col = factor)) +
    geom_line(alpha = 0.3) +
    geom_point(alpha = 0.8) +
    geom_line(data = y, aes(theta, yhat, group = NA, col = NA), alpha = 1)

```{r fig.cap = "\\label{fig:err} Errors"}
p1 <- ggplot(nonlg_est, aes(se, interval)) + geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)
p2 <- ggplot(nonlg_est, aes(se, diff)) + geom_point(col = "#56B4E9", alpha = 0.2)

grid.arrange(p1, p2, ncol = 2)

#### Executive Constraints Estimates

In total there are `r nrow(final.df)` country-year estimates of
$\theta$, comprised of `r n_distinct(final.df$country_name)` unique
countries and scaled from approximately $-2$ to $2$.  Since not every
single country can be presented, figure \ref{fig:theta} provides a
time series plot for China, Brazil, Nigeria, Norway, and Russia so as
to provide an example of the actual $\theta$ estimates.  The countries
were chosen to represent a diversity of geographic regions as well as
political contexts. From the plot, China, being a single party
authoritarian state, is consistently ranked the worst in executive
constraints. The decline for Russia largely follows the ascent of
Vladimir Putin at the end of 1999 and his gradual consolidation of
power in the decades that followed. For both Nigeria and Brazil, the
measure captures periods of military dictatorship such as the period
following Johnson Aguiyi-Ironsi's seizure of power in Nigeria the 1966
coup d'etat. The last country, Norway, represents a consolidated
polyarchy for the time period under consideration with consistently
strong constraints against the executive office.

```{r fig.width = 7, fig.height = 6, fig.cap = "\\label{fig:theta} Theta"}
theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)

theta_summarised <- theta_summarised[final.df$reduced_idx, ] %>%
    mutate(country_name = final.df$country_name,
           year = final.df$year)

targets <- c("Brazil", "China", "Nigeria", "Russia", "Norway")

filter(theta_summarised, country_name %in% targets) %>%
    ggplot(aes(year, median, fill = country_name)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2)

```{r fig.cap = "\\label{fig:tie} Tie Fighter"}
targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
             "Egypt", "Georgia", "Ghana", "Hungary", "Indonesia",
             "Iran", "Israel", "Japan", "Kenya", "Mexico", "Poland",
             "Portugal", "Romania", "Rwanda", "Turkey", "Tunisia",
             "United States of America", "Zimbabwe")

df <- filter(theta_summarised, year %in% c(2017, 2007),
             country_name %in% targets) %>%
    mutate(year = as.factor(year),
           country_name = ifelse(country_name == "United States of America",
                                 "USA", country_name) %>%
               factor(levels = sort(c(targets, "USA"), T)))

ggplot(df, aes(median, country_name, colour = year)) +
    geom_point() +
    geom_errorbarh(aes(xmax = codehigh, xmin = codelow))

```{r results = "hide"}
romania_idx <- filter(final.df, country_name == "Romania", year %in% c(2007, 2017)) %$%
    reduced_idx

romania_p <- signif(mean(theta[, romania_idx[1]] > theta[, romania_idx[2]]) * 100, 2)

chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2007, 2017)) %$%
    reduced_idx

chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
```

To provide a further snapshot of a greater number of countries, figure
\ref{fig:tie} plots the $\theta$ estimates for 2017, the last year in
the dataset, compared to 2007 by country. A full list of all countries
is available in [Appendix C](#appendix-c). Note, the statistical
significance of differences between years for a single country cannot
strictly be judged by overlapping credible intervals (TODO: Schenker
and Gentleman). For example, Romania, despite having overlapping
credible intervals, still has a posterior probability of approximately
$`r romania_p`\%$ that the estimate for 2017 is lower than 2007. In
the case of Chile with even more visual overlap, there is a
$`r chile_p`\%$ posterior probability that the estimate for 2017 is higher
than 2007.

### Onset Regression

Figure \ref{fig:reg} plots the slope coefficients on the log-odds
scale from the logistic regression linking the latent construct for
executive constraints to the outcome, civil conflict onset. Credible
intervals are summarised the same as previous plots, 80 percent
quantile intervals, and exact values are reported in table
\ref{tab:reg}.

- Log-odds scale
- Standardized

```{r reg, warning = F}
reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")

beta <- extract(reg, "beta")
colnames(beta) <- c("exec_constraints", colnames(X),
                    sprintf("exec * %s", colnames(X)[interaction_idx]))

probs <- apply(beta, 2, function(x) list(below = mean(x < 0), above = mean(x > 0)))

beta_s <- post_summarise(beta, probs = c(0.1, 0.9)) %>%
    mutate(Estimate = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3))) %>%
    select(-`10%`, -`90%`, -parameter)

rownames(beta_s) <- colnames(beta)

kable(beta_s, "latex", align = "c", booktabs = T, linesep = "", escape = T,
      caption = "Posterior credible intervals for regression slopes") %>%
    kable_styling(latex_options = c("striped", "scale_down"), full_width = F)
```

```{r fig.cap = "\\label{fig:reg} hello world"}
plot_pars(beta)
```

```{r eval = F}
# Interaction plot
alpha <- extract(reg, "alpha")
theta_seq <- seq(-2.5, 2.5, .01)

state <- quantile(final.df$rgdpepc, probs = c(0.1, .5, 0.9))

f <- function(x, y) {
    with(final.df, alpha +
                   beta[, 1] * x +
                   beta[, 2] * y +
                   beta[, 3] * mean(rgdpepc_gro) +
                   beta[, 4] * mean(pop_density) +
                   beta[, 5] * mean(meanelev) +
                   beta[, 6] * mean(rlvt_groups_count) +
                   beta[, 7] * 1 +
                   beta[, 8] * 1 +
                   beta[, 9] * mean(peace_yrs) +
                   beta[, 10] * 0 +
                   beta[, 11] * x * y)
}

par(mfrow = c(1, 3))
for (y in state) {
    print(y)
    p <- boot::inv.logit(sapply(theta_seq, f, y)) * 100

    plot(theta_seq, m, type = "l", ylim = c(min(pi[, 1]), max(pi[, 2])),
         xlab = "Executive Constraints", ylab = "Predicted Probability")
    lines(theta_seq, pi[, 1], lty = 2)
    lines(theta_seq, pi[, 2], lty = 2)
}
```

```{r eval = F}
input.df <- readRDS("posteriors/model/final.rds")
p_hat <- readRDS("posteriors/model/summary/p_hat.rds")

@@

```{r eval = F}
p <- apply(p_hat, 2, median)
x <- evalmod(scores = p, labels = input.df$lepisode_onset)

plot(x, curvetype = c("ROC", "PRC"))
```
